---
title: "HW2"
author: "Naumova Vika"
date: '17 мая 2018 г '
output: html_document
---
title
HW 2: From correlation to linear mixed-effect models. Assignment sheet
library(tidyverse)
library(lme4)
library(vcd)



1. Vowel reduction in Russian
Pavel Duryagin ran an experiment on perception of vowel reduction in Russian language. The dataset shva includes the following variables:
time1 - reaction time 1
duration - duration of the vowel in the stimuly (in milliseconds, ms)
time2 - reaction time 2
f1, f2, f3 - the 1st, 2nd and 3rd formant of the vowel measured in Hz (for a short introduction into formants, see here)
vowel - vowel classified according the 3-fold classification (A - a under stress, a - a/o as in the first syllable before the stressed one, y (stands for shva) - a/o as in the second etc. syllable before the stressed one or after the stressed syllable, cf. g[y]g[a]t[A]l[y] gogotala `guffawed').
In this part, we will ask you to analyse correlation between f1, f2, and duration. The dataset is available https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt.

1.0 Read the data from file to the variable shva.
```{r}
data <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt', sep = '\t')
```

1.1 Scatterplot f1 and f2 using ggplot().
Design it to look like the following.
```{r}
 ggplot(data, aes(f2, f1))+
  geom_point(aes(col = vowel))+
  scale_y_reverse()+
  scale_x_reverse()
```

1.2 Plot the boxplots of f1 and f2 for each vowel using ggplot().
Design it to look like this and this.
```{r}
ggplot(data, aes(vowel, f1))+
  geom_boxplot(aes(fill = vowel))+
  coord_flip()

ggplot(data, aes(vowel, f2))+
  geom_boxplot(aes(fill = vowel))+
  coord_flip()
```
  
# f1 boxplot
  

# f2 boxplot
1.3 Which f1 can be considered outliers in a vowel?
We assume outliers to be those observations that lie outside 1.5 * IQR, where IQR, the 'Inter Quartile Range', is the difference between the 1st and the 3rd quartile (= 25% and 75% percentile).
```{r}
boxplot(data$f1[data$vowel == 'a'])$out
``` 
1.4 Calculate Pearson's correlation of f1 and f2 (all data)
```{r}
cor.test(data$f1, data$f2)
```
1.5 Calculate Pearson's correlation of f1 and f2 for each vowel
```{r}
cor.test(data$f1[data$vowel == 'a'], data$f2[data$vowel == 'a'])
cor.test(data$f1[data$vowel == 'A'], data$f2[data$vowel == 'A'])
cor.test(data$f1[data$vowel == 'y'], data$f2[data$vowel == 'y'])
```
1.6
```{r}
model <- lm(f2 ~ f1, data)
```

1.6.1
```{r}
summary(model)$call
```
1.6.2
```{r}
summary(model)$adj.r.squared
```
1.6.3
```{r}
ggplot(data, aes(f2, f1))+
  geom_point(aes(col = vowel))+
  scale_y_reverse()+
  scale_x_reverse()+
  geom_smooth(method='lm', se = F, col = 'gray')
```
1.7
```{r}
mix_model <- lmer(f1 ~ f2 + (1|vowel), data)
```
1.7.1
```{r}
summary(mix_model)$call
```
1.7.2
```{r}
summary(mix_model)$varcor$vowel[1]
```
1.7.4
```{r}
data$mix_pred <- predict(mix_model)
ggplot(data, aes(x = f2, y = f1))+
  geom_point(aes(col = vowel))+
  scale_y_reverse()+
  scale_x_reverse()+
  geom_line(data = data, aes(x = f2, y = mix_pred, col = vowel), size = 0.7)
```
2. English Lexicon Project data
880 nouns, adjectives and verbs from the English Lexicon Project data (Balota et al. 2007).

Format -- A data frame with 880 observations on the following 5 variables.
Word -- a factor with lexical stimuli.
Length -- a numeric vector with word lengths.
SUBTLWF -- a numeric vector with frequencies in film subtitles.
POS -- a factor with levels JJ (adjective) NN (noun) VB (verb)
Mean_RT -- a numeric vector with mean reaction times in a lexical decision task Source (http://elexicon.wustl.edu/WordStart.asp)
Data from Natalya Levshina's RLing package available (here)[https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/ELP.csv]

2.0 Read the data from file to the variable elp.
```{r}
engdata <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/ELP.csv', sep = ',')
```

2.1 Which two variables have the highest Pearson's correlaton value?
```{r}
cor_matrix <- engdata %>% select_if(is.numeric) %>% cor
max_cor <- max(cor_matrix[lower.tri(cor_matrix)])
rownames(which(cor_matrix == max_cor, arr.ind=TRUE))
```

2.2 Group your data by parts of speech and make a scatterplot of SUBTLWF and Mean_RT.
```{r}
ggplot(engdata, aes(SUBTLWF, Mean_RT))+
  geom_point(aes(col = Length))+
  scale_color_continuous(low = "lightblue", high = "red")+
  facet_wrap(~ POS)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  coord_trans(x = "log10")
```
We've used scale_color_continuous(low = "lightblue", high = "red") as a parameter of ggplot().

2.3 Use the linear regression model to predict Mean_RT by log(SUBTLWF) and POS.
```{r}
model2 <- lm(Mean_RT ~ log(SUBTLWF) + POS, engdata)
```
2.3.1 Provide the result regression formula
```{r}
summary(model2)$call
```
2.3.2 Provide the adjusted R$^2$
```{r}
summary(model2)$adj.r.squared
```
2.3.3 Add the regression line in the scatterplot.
```{r}
ggplot(engdata, aes(log(SUBTLWF), Mean_RT))+
  geom_point(aes(col = Length))+
  scale_color_continuous(low = "lightblue", high = "red")+
  geom_smooth(method='lm', se = F, col = 'gray')
```

2.4 Use the mixed-efects model to predict Mean_RT by log(SUBTLWF) using POS intercept as a random effect
```{r}
mix_model2 <- lmer(Mean_RT ~ log(SUBTLWF) + (1|POS), engdata)
```
2.4.1 Provide the fixed effects formula
```{r}
summary(mix_model2)$call
```

2.4.2 Provide the variance for intercept argument for POS random effects
```{r}
summary(mix_model2)$varcor$POS[1]
```

2.4.3 Add the regression line to the scatterplot
```{r}
engdata$mix_pred <- predict(mix_model2)
ggplot(data2, aes(log(SUBTLWF), Mean_RT))+
  geom_point(aes(col = POS))+
  facet_wrap(~ POS)+
  geom_line(data = engdata, aes(x = log(SUBTLWF), y = mix_pred), col = 'black', size = 0.7)
```

3. Dutch causative constructions
This is a data set with examples of two Dutch periphrastic causatives extracted from newspaper corpora.

The data frame includes 100 observations on the following 7 variables:

Cx -- a factor with levels doen_V and laten_V
CrSem -- a factor that contains the semantic class of the Causer with levels Anim (animate) and Inanim (inanimate).
CeSem -- a factor that describes the semantic class of the Causee with levels Anim (animate) and Inanim (inanimate).
CdEv -- a factor that describes the semantic domain of the caused event expressed by the Effected Predicate. The levels are Ment (mental), Phys (physical) and Soc (social).
Neg -- a factor with levels No (absence of negation) and Yes (presence of negation).
Coref -- a factor with levels No (no coreferentiality) and Yes (coreferentiality).
Poss -- a factor with levels No (no overt expression of possession) Yes (overt expression of possession)
Data from Natalya Levshina's RLing package available (here)[https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv]

3.0 Read the data from file to the variable d_caus.
```{r}
caus <-  read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv')
```

3.1 We are going to test whether the association between Aux and other categorical variables (Aux ~ CrSem, Aux ~ CeSem, etc) is statistically significant. The assiciation with which variable should be analysed using Fisher's Exact Test and not using Pearson's Chi-squared Test? Is this association statistically significant?
```{r}
table(caus$CeSynt)
fisher.test(caus$Aux, caus$CeSynt)$p.value 
```

3.2. Test the hypothesis that Aux and EPTrans are not independent with the help of Pearson's Chi-squared Test.
```{r}
chisq.test(caus$Aux, caus$CrSem)$p.value
chisq.test(caus$Aux, caus$CeSem)$p.value 
chisq.test(caus$Aux, caus$CdEvSem)$p.value
chisq.test(caus$Aux, caus$EPTrans)$p.value 
chisq.test(caus$Aux, caus$Country)$p.value 
chisq.test(caus$Aux, caus$Domain)$p.value 
```
3.3 Provide expected values for Pearson's Chi-squared Test of Aux and EPTrans variables.
```{r}
(chi_exp <- chisq.test(caus$Aux, caus$EPTrans)$expected)
```
3.4. Calculate the odds ratio.
```{r}
(chi_exp[1][1]/chi_exp[3][1])/(chi_exp[3][1]/chi_exp[4][1])
```


3.5 Calculate effect size for this test using Cramer's V (phi).
```{r}
library(cramer)
cramer.test(caus$Aux, caus$EPTrans)
```
3.7
```{r}
library(vcd)
vcd::mosaic(~ Aux + CrSem + Country, data=d_caus, shade=TRUE, legend=TRUE)
```