---
title: "HW2"
author: "Naumova Vika"
date: '17 мая 2018 г '
output: html_document
---
title
HW 2: From correlation to linear mixed-effect models. Assignment sheet

library(tidyverse)
library(lme4)
library(vcd)


1.0 Read the data from file to the variable shva.
```{r}
data <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt', sep = '\t')
```

1.1 Scatterplot f1 and f2 using ggplot().
Design it to look like the following.
```{r}
 ggplot(data, aes(f2, f1))+
  geom_point(aes(col = vowel))+
  scale_y_reverse()+
  scale_x_reverse()
```

1.2 Plot the boxplots of f1 and f2 for each vowel using ggplot().
Design it to look like this and this.
```{r}
ggplot(data, aes(vowel, f1))+
  geom_boxplot(aes(fill = vowel))+
  coord_flip()

ggplot(data, aes(vowel, f2))+
  geom_boxplot(aes(fill = vowel))+
  coord_flip()
```
  
# f1 boxplot
  

# f2 boxplot
1.3 Which f1 can be considered outliers in a vowel?
We assume outliers to be those observations that lie outside 1.5 * IQR, where IQR, the 'Inter Quartile Range', is the difference between the 1st and the 3rd quartile (= 25% and 75% percentile).
```{r}
boxplot(data$f1[data$vowel == 'a'])$out
``` 
1.4 Calculate Pearson's correlation of f1 and f2 (all data)
```{r}
cor.test(data$f1, data$f2)
```
1.5 Calculate Pearson's correlation of f1 and f2 for each vowel
```{r}
cor.test(data$f1[data$vowel == 'a'], data$f2[data$vowel == 'a'])
cor.test(data$f1[data$vowel == 'A'], data$f2[data$vowel == 'A'])
cor.test(data$f1[data$vowel == 'y'], data$f2[data$vowel == 'y'])
```
1.6
```{r}
model <- lm(f2 ~ f1, data)
```

1.6.1
```{r}
summary(model)$call
```
1.6.2
```{r}
summary(model)$adj.r.squared
```
1.6.3
```{r}
ggplot(data, aes(f2, f1))+
  geom_point(aes(col = vowel))+
  scale_y_reverse()+
  scale_x_reverse()+
  geom_smooth(method='lm', se = F, col = 'gray')
```
1.7
```{r}
mix_model <- lmer(f1 ~ f2 + (1|vowel), data)
```
1.7.1
```{r}
summary(mix_model)$call
```
1.7.2
```{r}
summary(mix_model)$varcor$vowel[1]
```
1.7.4
```{r}
data$mix_pred <- predict(mix_model)
ggplot(data, aes(x = f2, y = f1))+
  geom_point(aes(col = vowel))+
  scale_y_reverse()+
  scale_x_reverse()+
  geom_line(data = data, aes(x = f2, y = mix_pred, col = vowel), size = 0.7)
```
2. English Lexicon Project data
880 nouns, adjectives and verbs from the English Lexicon Project data (Balota et al. 2007).


2.0 Read the data from file to the variable elp.
```{r}
engdata <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/ELP.csv', sep = ',')
```

2.1 Which two variables have the highest Pearson's correlaton value?
```{r}
cor_matrix <- engdata %>% select_if(is.numeric) %>% cor
max_cor <- max(cor_matrix[lower.tri(cor_matrix)])
rownames(which(cor_matrix == max_cor, arr.ind=TRUE))
```

2.2 Group your data by parts of speech and make a scatterplot of SUBTLWF and Mean_RT.
```{r}
ggplot(engdata, aes(SUBTLWF, Mean_RT))+
  geom_point(aes(col = Length))+
  scale_color_continuous(low = "lightblue", high = "red")+
  facet_wrap(~ POS)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  coord_trans(x = "log10")
```
We've used scale_color_continuous(low = "lightblue", high = "red") as a parameter of ggplot().

2.3 Use the linear regression model to predict Mean_RT by log(SUBTLWF) and POS.
```{r}
model2 <- lm(Mean_RT ~ log(SUBTLWF) + POS, engdata)
```
2.3.1 Provide the result regression formula
```{r}
summary(model2)$call
```
2.3.2 Provide the adjusted R$^2$
```{r}
summary(model2)$adj.r.squared
```
2.3.3 Add the regression line in the scatterplot.
```{r}
ggplot(engdata, aes(log(SUBTLWF), Mean_RT))+
  geom_point(aes(col = Length))+
  scale_color_continuous(low = "lightblue", high = "red")+
  geom_smooth(method='lm', se = F, col = 'gray')
```

2.4 Use the mixed-efects model to predict Mean_RT by log(SUBTLWF) using POS intercept as a random effect
```{r}
mix_model2 <- lmer(Mean_RT ~ log(SUBTLWF) + (1|POS), engdata)
```
2.4.1 Provide the fixed effects formula
```{r}
summary(mix_model2)$call
```

2.4.2 Provide the variance for intercept argument for POS random effects
```{r}
summary(mix_model2)$varcor$POS[1]
```

2.4.3 Add the regression line to the scatterplot
```{r}
engdata$mix_pred <- predict(mix_model2)
ggplot(data2, aes(log(SUBTLWF), Mean_RT))+
  geom_point(aes(col = POS))+
  facet_wrap(~ POS)+
  geom_line(data = engdata, aes(x = log(SUBTLWF), y = mix_pred), col = 'black', size = 0.7)
```

3. Dutch causative constructions


3.0 Read the data from file to the variable d_caus.
```{r}
caus <-  read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv')
```

3.1 We are going to test whether the association between Aux and other categorical variables (Aux ~ CrSem, Aux ~ CeSem, etc) is statistically significant. The assiciation with which variable should be analysed using Fisher's Exact Test and not using Pearson's Chi-squared Test? Is this association statistically significant?
```{r}
table(caus$CeSynt)
fisher.test(caus$Aux, caus$CeSynt)$p.value 
```

3.2. Test the hypothesis that Aux and EPTrans are not independent with the help of Pearson's Chi-squared Test.
```{r}
chisq.test(caus$Aux, caus$CrSem)$p.value
chisq.test(caus$Aux, caus$CeSem)$p.value 
chisq.test(caus$Aux, caus$CdEvSem)$p.value
chisq.test(caus$Aux, caus$EPTrans)$p.value 
chisq.test(caus$Aux, caus$Country)$p.value 
chisq.test(caus$Aux, caus$Domain)$p.value 
```
3.3 Provide expected values for Pearson's Chi-squared Test of Aux and EPTrans variables.
```{r}
(chi_exp <- chisq.test(caus$Aux, caus$EPTrans)$expected)
```
3.4. Calculate the odds ratio.
```{r}
(chi_exp[1][1]/chi_exp[3][1])/(chi_exp[3][1]/chi_exp[4][1])
```


3.5 Calculate effect size for this test using Cramer's V (phi).
```{r}
library(cramer)
cramer.test(caus$Aux, caus$EPTrans)
```
3.7
```{r}
library(vcd)
vcd::mosaic(~ Aux + CrSem + Country, data=d_caus, shade=TRUE, legend=TRUE)
```